\documentclass[11pt,a4paper,oneside]{article}
\renewcommand{\baselinestretch}{1.2}
\usepackage[top=1.00in, bottom=1.0in, left=1in, right=1.25in]{geometry} 
\usepackage{graphicx}
\usepackage{latexsym,amssymb,epsf} 
\usepackage{epstopdf}
\usepackage{amsmath}
\usepackage{hyperref}

% Optional, you can get rid of this:
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhead[LO]{Leftside header here}
\fancyhead[RO]{Rightside header here}

\begin{document}

You can put inline math in with blackslash-parenthesis, then math, then blackslash-parenthesis, like this: \(\psi+\phi\).\\

For longer series you can use align (add an asterisk to remove automatic numbering of equations) and the \& sign (saying where you want the alignment to happen, e.g., the equals sign or such):

\begin{align}
x_{i} =& y+g^{\theta} \text{   :basic fun} 
\\
s_{i} =& x+i
\end{align}

\noindent Handy websites:\\
\href {http://en.wikibooks.org/wiki/LaTeX/Mathematics}{Latex Wikibooks: Math}\\
\href{http://en.wikibooks.org/wiki/LaTeX/Advanced_Mathematics}{Latex Wikibooks: Advanced Math}\\
\href{http://detexify.kirelabs.org/classify.html}{Draw your symbol in order to find it!}

\noindent \bf{Page 1}\\
\noindent Chesson 2003, Eqn 1
\begin{align}
N_{j}(t+1) = (1-d_{j}) N_{j} + R_{t}N_{j}
\end{align}
where 
\\d_{j} is \emph{per capita} death of persistent stage (for us, seeds)
\\and
\\R_{j} is \emph{per capita} number of new recruits

This is the general form of the model.  When translating our model (from Chesson 2004) into these terms, we must decide how to measure survivorship (i.e., \((1-d_{j})).  It can be either \(s_{j}) or \(s_{j}(1-g_{j}(t)).
\\The first case is cleaner -- each species has a characteristic survivorship and all the germination dynamics are in the recruitment function.  But survivorship is used later to scale for generation time.  In that case, what is the appropriate generation time?  

The equivalent in our model is 
\begin{align}
N_{j}(t+1) = s_{j}N_{j}(t)(1-g_{j}(t)) + s_{j}N_{j}(t)g_{j}(t)\phi_{j}B_{j}(t+\delta)
N_{j}(t+1) = \text{survival of persistent stage and loss to germination} + \text{recruitment to persistent stage from germinants}
=s_{j}N_{j}(t) + N_{j} + R_{t}N_{j}
= \text{survivorship of persistent stage} + \text{recruitment to persistent stage}
\\ \text{where}
R_{t} = g_{j}(t)s_{j}N_{j}(t)[\phi B_{j}(t+\delta)-1]
\\ R_{t}\text is now germination, growth, conversion back to seeds, and overwinter survival 

Equivalents in our model (from Chesson 2004)
\end{document}

% Handy commands %
\noindent 
\section{name}
\subsection{name}
\emph{italics!} 
{\bf for bold}
\newpage 

% Options to fancy up first page (but these after begin{document})
\title{}
    \author{moi}
\date{\today}
\maketitle

\begin{abstract} 
\end{abstract}