\documentclass[11pt]{article}
\usepackage[top=1.00in, bottom=1.0in, left=1.1in, right=1.1in]{geometry}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{lineno}
\usepackage{xr-hyper}
\externaldocument{..//..//..//phencc}
\newcommand{\lr}[1]{line~\lineref{#1}}
% \usepackage{hyperref}
\setlength\parindent{0pt}

% Need to provide:
% As such, pleaes could I ask you to prepare a revision, including a point-by-point response to the reviewers' constructive suggestions?
% Revise your manuscript using a word processing program and save it on your computer.  Please also highlight the changes to your manuscript within the document by using the track changes mode in MS Word or by using bold or colored text.
% When submitting your revised manuscript, you will be able to respond to the comments made by the reviewer(s) in the space provided.

\begin{document}
Editor and reviewer comments are in \emph{italics} (we provide below the full context of each review), while our responses are in regular text. \\ 

{\bf Editor's comments:} \\

\emph{As you can see from the comments of the expert reviewers below, both are broadly positive
about the manuscript, but both have suggestions for further improvement. Referee 1, in
particular, thinks that the submission would benefit from some fairly substantial rewriting
in places, and I would tend to agree.}\\

We thank the editor for the opportunity to revise our manuscript. We found the reviewers' comments very helpful, with many overlapping requests for clarity or a more careful message. We provide detailed point-by-point responses below.\\

{\bf Reviewer 1 comments:} \\


\emph{The authors suggest that tracking individual species' phenological response to environmental
change may be insufficient to meaningfully predict community response to future climate
change due to the influence of species' interdependence beyond trophic interactions and other
influences at the community level. In brief, whether or not a particular species will survive
climate change is complex and depends on a range of external factors that affect fitness
including the influence of other species in a dynamic system that cannot be predicted by
simply tracking its phenology. The impact of climate change on species' interactions with the
physical environment and other organisms is notoriously complex and difficult to disentangle.
Overall, this is a very timely and important topic and I would support its publication after
tightening up of the language and presenting a sharper, clearer message. As it stands this
manuscript appears wordy and sometimes laboured in places which tends to mask the key message
being conveyed. It would be useful to highlight how this manuscript contributes to advancing
the field of predicting species' response to climate change and how this in turn may impact
community assemblages.}\\

Thanks for this very helpful review. We completely agree that we could make the message clearer and make the writing easier on readers. As such, we have made changes throughout, from the abstract to the conclusions. In particular we have tried to reduce the wordiness, including the asides, caveats and examples. We have a lot of these! Partly they come from the struggle of trying to unite the empirical and theoretical framework: the theorist on the author team aims for precision, which led to some of these additions, while the empiricist wanted to clarify what we meant for other empiricists, which led to more of these additions. We have worked to remove many of them, leaving the ones we think are most important for comprehension (but do let us know if you see more that could be removed).  \\

\emph{ Abstract\\
Consider replacing `of' with `for' in the first sentence. Consider removing `us' in second
last sentence.}\\

We changed the first sentence a little (we stuck with `of' as we found `for' read odd given other changes to the sentence), and we removed `us.' Also, we re-wrote the abstract a good bit for clarity and to better state the aim of our review. \\

\emph{Briefly define `stationary and non-stationary systems'. Consider conveying in this sentence
that the proposed method will `help' predict rather than definitively `predict'. It would be
useful to emphasize the key findings of the review.}\\

We re-wrote the abstract substantially to emphasize our aims and findngs. Also, we have changed to `help understand.' Given our definition of stationary and non-stationary we were not able to briefly define these terms in the abstract, however, we have added a glossary to the paper as the first table. \\

\emph{Introduction\\
In the second sentence, consider mentioning phenology before space to be consistent with
`time and space' at the end of the first sentence.}\\

Good catch! We switched the ordering of `space and time' in this first sentence, and altered the second sentence to make the connection to `time' in the first sentence more clear.\\

\emph{P4L12 Consider adding `and locations' after `..across species'.}\\

Done.\\

\emph{P4L38 It would be useful to clarify what is meant by `...the underlying distribution of the
environment.....' –is this related physical climatic parameters? }\\

Good point, we have done this in the new glossary that we added.\\

\emph{P4L53/54 Consider replacing `measuring tracking in current environments and evaluating the
fitness outcomes of tracking' with `measuring tracking and evaluating its fitness outcomes,
in current environments'. }\\

Done. We also just removed `in current environments' to make the sentence less wordy.\\

\emph{In Figure 1 it would be useful to know how many years before and after 1980 were included in
the analysis. }\\

We hinted at this at this at the end of the caption, but can see now it was not very clear. We have 40 years before 1980 and 40 years after 1980 for all the datasets, which we now say early in the caption. \\

\emph{What is the specific aim of this review – the authors mention what they are going to review
but why is it necessary and what is this synthesis expected to reveal? }\\

Great point, we have added this here (paragraph starting on \lr{r1whatpoint}) and in the abstract. We also adjusted our conclusions some to address this. \\

\emph{Defining and measuring tracking\\
Consider sticking to vegetation or animal examples not a combination of both or explain the
rationale for using both groups.}\\

We have restructured this section slightly and now address this question in \lr{r1whyanimals}-\lr{r1whyanimalsend}.\\

\emph{L42 consider replacing `these' with `the timing and/or intensity'. }\\

We have removed this sentence (we believe, we're not sure of page number, let us know if not).\\

\emph{Overall, I found section 2 very wordy and difficult to follow. P5L47 Consider the following
sentence `Tracking is commonly used to describe how phenology responds to climate change, yet
it is rarely defined'. This section could be clearer and `tighter'.}\\

Agreed! We made lots of changes here, including the one suggested.\\

\emph{P6L25/26 consider replacing `....to be something that can be accurately modeled..' with `to
permit/allow accurate modeling'.}\\

Done.\\

\emph{P6L33/34 remove `a' before `interaction'.}\\

We believe this text has been removed in the revision process.\\

\emph{Does environmental tracking include parameters other than climatic variables and photoperiod,
such as, nutrient availability etc.? It might be useful to state this.}\\

Environmental tracking is a product of an organism's cue system, and we have worked to clarify this in the text, especially in the glossary. To date major cue systems have found evidence of photoperiod, temperature, moisture cues mainly. Based on this we expect that environmental tracking is not directly related to nutrients in many systems. However, it may be correlated with tracking, as the cue system is tuned to maximizing nutrient availability (but is not responding directly to it).  \\

\emph{Another possible reason that species may not appear to `track' maybe that some species
require a greater amount of change in the environmental cue i.e. a higher threshold, before they respond. It seems likely that there are more than 3 major reasons why tracking is not detectable. A clear rationale for selecting and focusing on these reasons would be useful.}\\

Good point, we actually see that reason as covered by our first reason, but we agree that there could be others and have adjusted our text here somewhat. \\

\emph{P7L40/41 consider placing `briefly' before `review'.}\\

Done.\\

\emph{P9L18/19 replace `that' with `the'.}\\

Done.\\

\emph{P10L54/55 repetition of `fluctuations in the environment' consider something like `mechanisms
which are dependent on, or independent of, fluctuations in the environment.....}\\

Done.\\

\emph{P11L9/10 Consider `In community ecological modeling, definitions of the environment generally
fall into two broad categories'.}\\

Done.\\

\emph{P13L5/6 consider removing `including the previous example'. L16-18 is it necessary to be so
explicit? Throughout the MS I find much of the information presented within brackets
distracting. }\\

Removed and we agree! We have tried to remove these caveats and `let's be super explicit' additions throughout. We were trying to be exact and thorough, but it makes it hard to read and isn't always needed.\\

\emph{P13L21-25 if the resource is in limited supply this may provide an advantage to
early arrivals but later arrivals will benefit from the same resource if it is not depleted
by earlier arrivals and other conditions may be better such as less risk of late frost.
It is well established that the environmental variables being monitored are not exactly what
the organism is responding to as micro-climate varies considerably in even the simplest
ecosystem, furthermore, do we even know if budburst for example is triggered to the
temperature of the bud surface, some internal temperature or some interaction with root (or
some other organ) temperature. Defining the specific environmental variable and the specific
threshold each species and each phenophase is responding to appears overly complex – there
must be some trade-off between a researchers effort and the applicability of the results. I
just wonder where we draw (or don't) the line. This is a very challenging and important topic
that the authors are addressing.}\\

We agree that this is a very challenging topic. Where to draw the line is an important question and we hope our review highlights this complexity. We believe that different approaches must draw lines at different places. Each approach, however, likely benefits from a more explicit consideration of what line they draw and what lines other approaches draw, which we review---from the molecular level (extremely specific models such as the photo-thermal model of flowering of  \emph{Arabidopsisis thaliana}, now mentioned on \lr{r2arab}) to models of interspecific competition. \\

\emph{Conclusions are more of a summary of the topics reviewed rather than a comprehensive
synthesis of the literature to draw new and more advanced conclusions based on the collective
information from the review. It might be useful to make some recommendations on what is
needed and why. Therefore, perhaps point 5 could be expanded.}\\

We have expanded point 5 (now point 6) and adjusted our conclusions to more clearly state the path forward. Our sections 4.5 (`Frontiers of community assembly models') and section 5 (`Linking empirical and theoretical research') are focused on recommending what is needed, so we worked to summarize these better in the conclusions, while also following journal guidelines (which state that the ``conclusions section should be in the form of a short list of numbered points summarising the main findings of your article. Avoid the introduction of new material or new references in the conclusions section.'').  \\

{\bf Reviewer 2 comments:} \\

\emph{In this review the authors argue that phenological tracking data needs to be combined with
coexistence theory to help make predictions about how climate change will affect communities,
particularly in non-stationary environments. Generally, this was well-written, clear in its
logic, and really laid out how phenological tracking vs. coexistence theory are currently
divided but could be integrated. The one aspect that fell short was suggesting specific types
of experimental data that might help us reach this integration of phenological cues and
climate change in a multi-species framework. I also have a couple of clarifying questions
throughout. But otherwise, this was a really interesting review proposing a novel combination
of two research areas that would really strengthen predictions about the consequences of
climate change.}\\

We are excited to hear the reviewer found our manuscript useful. We hope our revision kept the good aspects and improved the clarity of the message overall.\\

\emph{1.      The argument for phenological tracking combined with coexistence theory in stationary
environments itself is something that needs more work, but it was kind of glossed over in
favor of non-stationary environments. It might be worth laying out more about the gaps
between tracking and coexistence in stationary environments in the intro (pg.4, between lines
33 and 34?).}\\

We tried to address this as best we could without impacting the flow and coherence of the introduction (see \lr{r2stat}-\lr{r2statend}). As we are focused here on climate change and thus non-stationarity our introduction is focused on these interwoven topics.\\

\emph{2.      Pg.4, L49: Can you be more specific about examples of the community-level processes
here? At first I thought this meant phenological tracking as a process but that would be
individual or population level.}\\

Yes! We have added `such as competition and priority effects' (see line \lr{r2example}).\\

\emph{3.      Pg.5, L31: Isn't number of offspring still measured at the individual level? What
does "higher levels" mean here? Perhaps an example like synchrony at the community level
would fit better?}\\

Good point, switched to population.\\

\emph{4.      Pg.5, L35: I agree that the first event is not equivalent to the number of flowers
(or a continuous phenological metric), but yes/no flowering is not the same as first event
either – these variables seem conflated in this section.}\\

We have removed the sentence about first events.\\

\emph{5.      Pg.6, L34-36: This section made me wonder if there are any examples where researchers
know the exact cues and can measure phenological tracking more precisely in systems like
Arabidopsis?}\\

Great point! We have added mention of the photothermal model of flowering of  \emph{Arabidopsisis thaliana} (\lr{r2arab}) and we also discuss it in a new section `Building from cue systems to phenological tracking' in our future directions (section 5, `Linking empirical and theoretical research), see \lr{r2hertiable}-\lr{r2hertiableend}.\\

\emph{6.      Pg.7, L49: This is somewhere where I thought empirical directions could be suggested
(or using this, have a section on future research at the end) – the heritability of
phenological tracking made me think maybe we need more studies linking phenological
plasticity to genetic variation and heritability (e.g., do more plastic populations harbor
more genetic variation in phenological traits)?}\\

Good point, we added this to \lr{r2hertiable}-\lr{r2hertiableend} in a new section `Building from cue systems to phenological tracking' in our future directions (section 5, `Linking empirical and theoretical research).\\

\emph{7.      Pg.8, L23: This is the other spot where I wanted more background/discussion of gaps
in the links between phenological tracking and stationary environments before moving on to
non-stationary environments.}\\

Good point, we have expanded on this, see \lr{r2wantsmore}-\lr{r2wantsmoreend}.\\

\emph{8.      Pg.8, L24: Is ``unpredictable'' environments here different from non-stationary
environments? In pg.9, L3, it sounds like non-stationary also means unreliable in this
context, so I was a bit unclear on why unpredictability was introduced in 3.1 instead of 3.2
(I might have missed some key difference). A glossary box for some of these terms could also
be helpful.}\\

Good point, this was rather inexact, we removed this word and are more specific now (see \lr{r2unpredict}).\\

\emph{9.      Pg.9, L18: What is the fundamental ``model'' here? Is this underlying fitness?}\\

Good point, we meant ideal timing (relating to the underlying model of fundamental tracking), which we now say explicitly (see \lr{r2whatmodel}).\\

\emph{10.     Pg. 9, L33: The idea that phenological tracking might vary not only in stationary vs.
non-stationary environments but also in environments transitioning from
stationary/predictable to non-stationary/non-predictable is cool (and brought up again later
in the review), but it's introduced so late in this section and without any context!
Definitely worth expanding on.}\\

Good point, we introduced it late as it is setup for our next section, but it does seem a big point that we're making rather too easy to miss. We have expanded to a full paragraph (see \lr{r2expand}-\lr{r2expandend}).\\

\emph{11.     Pg.9, L47: This statement felt short – perhaps include some reasons why competition
is critical. Additionally, this is another spot with some future directions implications
because phenological shifts aren't often studied in the context of their effects on species
interactions (although we often make assumptions based on biology, e.g. invasive-native
comparisons). Perhaps a call to include phenological shifts across environments as part of
per-capita competition experiments?}\\

Good point, we changed to `Yet decades of research show that competition drives the niche differences necessary for species to co-exist \citep{Hutchinson:1959xi,Chesson:2000vd}' (\lr{r2whycomp}). Also, we added this to our future directions (section 5, `Linking empirical and theoretical research), see \lr{r2exp}-\lr{r2expend}.\\

\emph{12.     Pg.10, first paragraph: I found this argument really interesting because although it
made sense, it read as counterintuitive to invasive-native comparisons of phenology. The
argument here is that more plastic species are probably inferior resource competitors. But
studies that find that invasive species are more plastic also assume invasive species are the
superior resource competitors. I wonder if this section needs some caveat in our assumptions
for invasive species in these models? (Similarly in 4.2, but I think that the argument that
``best-matched'' species could drive the other extinct (Pg.12, L45) starts to help with this.)}\\

This is a super interesting point, which we think depends on how you define superior and inferior resource competition. The superior competitor is often defined as the species that can draw down the resource to a lower level, and in the systems we have worked in this is often the native species (for example, invasive annual grasses versus native shrubs in coastal sage scrub in California or invasive shrub glossy buckthorn versus native trees in the US midwest). However, by being earlier the invasive species may outcompete the native through early access to resources, but it would not generally be defined as the superior competitor. Testing this, however, requires measuring more traits across native and invasive species (especially the difficult ones related to resource competition), which we now state in our conclusions (\lr{r2invas}).\\

\emph{13.     Pg.12, L55: What if phenological tracking itself could be a stabilizing mechanism by
affecting resource partitioning?}\\

This is a great question. Based on our current understanding of community assembly, tracking could only be stabilizing if it trades off with another trait or in some other way creates a niche difference. If tracking somehow partitions resources this could work, but it would have to partition across multiple resources or in some other way limit competition for the same resource across multiple species. \\

\emph{14.     Pg.13, L5-12: The review draws on plasticity literature, but there are definitely
studies linking plasticity (including phenological plasticity) to fitness that seem
overlooked in this first paragraph. E.g., Lustenhouwer, N., et al. (2018). Global Change
Biology 24(2): e534-e544.}\\

Agreed, we now address this in \lr{r2hertiable}-\lr{r2hertiableend} in a new section `Building from cue systems to phenological tracking.' This point in general is a very good one and something we have definitely struggled with in writing this manuscript. While we aim to focus on multi-species model, the literature on plasticity is critical to phenology. We have tried to include the major background from this very large literature as much as possible without greatly extending the length and focus of the review. \\

\emph{15.     Pg.13, L35: I'm unclear on whether phenology is explicitly in these models.
Otherwise, how is phenology connected to studies of production that don't measure phenology?}\\

We did not make this clear at all and have revised the text (\lr{r2thisrelates} - \lr{r2thisrelatesend}) to clarify. \\

\emph{16.     Pg.13, L38: Figure 6 is your model, it doesn't show whether other studies model
environmental distributions. Why referenced here?}\\

Good point! We have now switched to referencing several major papers on this (\lr{r2refs}).\\

\emph{17.     Pg.14, L25: Could you expand on why the new communities would exist only as long as
the environment remained non-stationary? Could plastic species that arrived in a non-
stationary environment also stay if the environment was just consistently warmer, for example
(i.e., stationary)?}\\

\emph{18.     Pg. 14, L48: I love this call to action – great place to put the types of data we
still need.}\\

Thank you!\\

\emph{19.     The figures are all great – I found 2 and 5 particularly helpful for some of the
complicated concepts in here. Figure 5a also made me think about whether we need more studies
changing the timing of resource (or stress) pulses and how that affects tracking.}\\

It's great to hear these figures are useful to others---as they have been useful to us. Studies of how changing the timing of resource pulses affects tracking would be very interesting, especially if they could be done manipulatively. We hope work inspired by our paper might provide the theory to best design such experiments. \\

\bibliography{/Users/Lizzie/Documents/git/bibtex/LizzieMainMinimal}
\bibliographystyle{/Users/Lizzie/Documents/git/bibtex/styles/ecolett.bst}

\end{document}
